Master thesis

### 2017-07-19 00-Abstract.Rmd
Abstract 1

Écrire le résumé français ici...

Write the english abstract here...

>>>>

Les forêts tropicales font face à de nombreuses perturbations qui représentent la troisième source mondiale d'émission de gaz à effet de serre. La déforestation et la dégradation des forêts tropicales sont responsables de l'émission de 8.26 milliards de tonnes de dioxyde de carbone par an [@Pearson2017]. La déforestation a retenu l'attention mondiale, mais la dégradation des forêts représente 20% des émissions de l'Amazonie brésilienne [@Asner2005]. La gestion durable des forêts a été proposée comme réponse à la déforestation et la dégradation, malgré la remise en question de la durabilité de l'exploitation forestière [@Zimmerman2012]. D'autre part, les forêts tropicales abritent plus de la moitié de la biodiversité terrestre mondiale [@Scheffers2012]. Par conséquent, nous avons décidé d'étudier le rôle de la biodiversité dans la réponse des écosystèmes forestiers aux perturbations, en reliant diversité et fonctionnement de l'écosystème [@Loreau2010]. Nous avons utilisé l'hypothèse que lors d'une perturbation, grâce à une productivité plus forte, une forêt plus diverse aura une meilleure résilience, en se basant sur la relation positive entre biodiversité et productivité. Nous avons relié cette hypothèse aux effets de complémentarité et de sélection [@Loreau2001]. La complémentarité est la combinaison de la partition des ressources et de la facilitation, alors que l'effet de sélection est le résultat de la sélection compétitive. Nous avons ainsi centré l'étude sur les mécanismes impliqués dans la relation entre biodiversité et résilience des écosystème forestiers par une approche par simulation afin d'appréhender les processus à long terme. Nous avons utilisé le modèle TROLL [@Li] pour simuler 60 forêts matures aux diversités taxonomiques et fonctionnelles croissantes. Nous avons perturbé toutes les forêts et mesuré la résilience de leurs fonctions écosystémiques. En outre, nous avons mesuré la résilience de l'effet net de la biodiversité que l'on a décomposé en en effets de complémentarité et de sélection. Nous avons trouvé que la diversité améliore la résilience des forêts tropicales, particulièrement au travers de la diversité et l'équitabilité fonctionnelle. De plus, nous avons montré que la complémentarité entre les espèces assurait la résilience de la forêt en début de succession avant de laisser place à l'effet de sélection. Nos résultats suggèrent la possibilité d'une gestion durable des forêts tropicales grâce à une meilleure résilience avec une plus haute diversité. Mais cette conclusion n'a de sens que si l'exploitation sélective est durable [@Zimmerman2012]. Au contraire, une gestion non durable des forêts tropicales entraînera des rétroactions négatives diminuant lentement la diversité et donc la résilience des forêts, aboutissant ultimement à la dégradation des forêts.

Forest disturbances are the third worldwide source of greenhous gas. Tropical deforestation and degradation emit 8.26 billion of tons of carbon dioxyde per year [@Pearson2017]. Deforestation has retained much attention, but degradation from forest represents 20% of emissions in brazilian Amazon [@Asner2005]. Sustainable forest management has been promoted as an answer to deforestation and degradation, besides logging sustainability has been questionned [@Zimmerman2012]. On the other hand, tropical forest host over half of the Earth's biodiversity [@Scheffers2012]. Consequently, we decided to study the role of biodiversity in forest ecosystem answer to disturbance, linking diversity to ecosystem functionning [@Loreau2010]. We used the hypothesis that when a disturbance event happen, due to a higher productivity, a more diverse forest will be more resilient, based on the positive relationship between biodiversity and productivity. We linked that hypothesis to the complementarity and selection effects [@Loreau2001]. Complementarity is the addition of ressource partitionning and facilitation, whereas selection effect is the result of competitive selection. We thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience with a simulation approach to assess long term processes. We used TROLL model [@Li] to simulate 60 matures forests with growing taxonomic and functional  diversities. We disturbed all forests and measured the resilience of their  ecosystem functions. Additionnally, we measured biodiversity net effect resilience partitioned into complementarity and selection effects. We found that diversity improved tropical forest resilience, particularly through functional diversity and eveness. Moreover, we showed that complementarity between species insured forest recovery in the beginning of the succession before being replaced by selection effect. Our results suggest the possibility for a sustainable management of tropical forest due to an increased resilience with an higher diversity. But this conclusion has meaning only if selective logging meet sustainability [@Zimmerman2012]. On the contrary, unsustainable tropical forest management will lead to negative feedbacks slowly diminishing diversity and thus forest resilience, resulting ultimately in forest degradation.

### 2017-07-19 02-Introduction.Rmd
Abstract 1

Tropical forests perturbations, through deforestation and degradation, account for 8.26 billion of carbon dioxyde emissions per year [@Pearson2017]. It represents the third source of greenhouse gas. Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics. Degradation from forest has been estimated to represents 10 time deforestation [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005]. Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.

@@

We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance highlighting tropical forest carbone resilience to disturbance. Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.

>>>>

Tropical forests disturbances, through deforestation and degradation, account for 8.26 billion of tons of carbon dioxyde emissions per year [@Pearson2017]. It represents the third source of greenhouse gas. Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics. Degradation from forest has been estimated to represents 10 time deforestation [@Herold2011] and represents 20% of greenhouse gas emissions in the brazilian Amazon [@Asner2005]. Degradation has met numerous definitions [@Simula2009], but can be defined as the result from a disturbance event that induced a modification of the forest ecosystem reducing ecosystem services while keeping the ecosystem as a forest, contrary to deforestation.

@@

We decided to use the forest model TROLL to study the role of biodiversity in forest ecosystem answer to disturbance. Resilience encompass several definitions but was summarized by @Oliver2015 as the degree of resistance or fast recovery from an ecosystem function to environmental disturbance.

### 2017-07-19 06-Discussion.Rmd
Abstract 1

In the limit of the model, we were able to show that diversity improved tropical forest resilience. More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance. Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity. But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012. Because if the harvesting is not sustainable negative feedbaks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.

>>>>

In the limit of the model, we were able to show that diversity improved tropical forest resilience. More particularly, functional diversity and evenness are key components of diversity in forest recovery after disturbance. Moreover, we found that complementarity between species was insuring forest recovery in forest succession start with facilitation before more productive species dominate the forest and insure recovery. Our results advocates for a sustainable harvesting of tropical forests through an increased resilience due to high diversity. But this conclusion should met a sustainable definiton of selective logging following @Zimmerman2012. Because if the harvesting is not sustainable negative feedbacks will slowly diminish diversity and its benefits for forest resilience, resulting in forest degradation.

### 2017-07-19 docs/master-thesis.tex
Abstract 1

Écrire le résumé français ici\ldots{}

Write the english abstract here\ldots{}

@@

\begin{itemize}
\tightlist
\item
  Bruno
\item
  Stéphane
\item
  Laurent
\item
  Fabian
\item
  Isabelle
\item
  Jérôme
\item
  Camilla
\item
  Aurélie
\item
  Éric
\item
  Maxime
\item
  Raphaël
\item
  Paracou team
\item
  \ldots{}
\end{itemize}

@@

Tropical forests perturbations, through deforestation and degradation,
account for 8.26 billion of carbon dioxyde emissions per year
\citep{Pearson2017}. It represents the third source of greenhouse gas.
Besides deforestation has been rightly the focus of worldwide attention,
degradation has been less studied and quantified through tropics.
Degradation from forest has been estimated to represents 10 time
deforestation \citep{Herold2011} and represents 20\% of greenhouse gas
emissions in the brazilian Amazon \citep{Asner2005}. Degradation has met
numerous definitions \citep{Simula2009}, but can be defined as the
result from a disturbance event that induced a modification of the
forest ecosystem reducing ecosystem services while keeping the ecosystem
as a forest, contrary to deforestation.

@@

We decided to use the forest model TROLL to study the role of
biodiversity in forest ecosystem answer to disturbance highlighting
tropical forest carbone resilience to disturbance. Resilience encompass
several definitions but was summarized by \citet{Oliver2015} as the
degree of resistance or fast recovery from an ecosystem function to
environmental disturbance.

@@

In the limit of the model, we were able to show that diversity improved
tropical forest resilience. More particularly, functional diversity and
evenness are key components of diversity in forest recovery after
disturbance. Moreover, we found that complementarity between species was
insuring forest recovery in forest succession start with facilitation
before more productive species dominate the forest and insure recovery.
Our results advocates for a sustainable harvesting of tropical forests
through an increased resilience due to high diversity. But this
conclusion should met a sustainable definiton of selective logging
following \citet{Zimmerman2012}. Because if the harvesting is not
sustainable negative feedbaks will slowly diminish diversity and its
benefits for forest resilience, resulting in forest degradation.

@@

% Last pages
  %Last page
  \newpage
  \paragraph{Résumé :}
  Écrire le résumé ici...
  \paragraph{Mots clés :} mots clés
  \newline\newline
  \paragraph{Abstract:}
  Write abstract here
  \paragraph{Keywords:} keywords

>>>>

Les forêts tropicales font face à de nombreuses perturbations qui
représentent la troisième source mondiale d'émission de gaz à effet de
serre. La déforestation et la dégradation des forêts tropicales sont
responsables de l'émission de 8.26 milliards de tonnes de dioxyde de
carbone par an \citep{Pearson2017}. La déforestation a retenu
l'attention mondiale, mais la dégradation des forêts représente 20\% des
émissions de l'Amazonie brésilienne \citep{Asner2005}. La gestion
durable des forêts a été proposée comme réponse à la déforestation et la
dégradation, malgré la remise en question de la durabilité de
l'exploitation forestière \citep{Zimmerman2012}. D'autre part, les
forêts tropicales abritent plus de la moitié de la biodiversité
terrestre mondiale \citep{Scheffers2012}. Par conséquent, nous avons
décidé d'étudier le rôle de la biodiversité dans la réponse des
écosystèmes forestiers aux perturbations, en reliant diversité et
fonctionnement de l'écosystème \citep{Loreau2010}. Nous avons utilisé
l'hypothèse que lors d'une perturbation, grâce à une productivité plus
forte, une forêt plus diverse aura une meilleure résilience, en se
basant sur la relation positive entre biodiversité et productivité. Nous
avons relié cette hypothèse aux effets de complémentarité et de
sélection \citep{Loreau2001}. La complémentarité est la combinaison de
la partition des ressources et de la facilitation, alors que l'effet de
sélection est le résultat de la sélection compétitive. Nous avons ainsi
centré l'étude sur les mécanismes impliqués dans la relation entre
biodiversité et résilience des écosystème forestiers par une approche
par simulation afin d'appréhender les processus à long terme. Nous avons
utilisé le modèle TROLL \citep{Li} pour simuler 60 forêts matures aux
diversités taxonomiques et fonctionnelles croissantes. Nous avons
perturbé toutes les forêts et mesuré la résilience de leurs fonctions
écosystémiques. En outre, nous avons mesuré la résilience de l'effet net
de la biodiversité que l'on a décomposé en en effets de complémentarité
et de sélection. Nous avons trouvé que la diversité améliore la
résilience des forêts tropicales, particulièrement au travers de la
diversité et l'équitabilité fonctionnelle. De plus, nous avons montré
que la complémentarité entre les espèces assurait la résilience de la
forêt en début de succession avant de laisser place à l'effet de
sélection. Nos résultats suggèrent la possibilité d'une gestion durable
des forêts tropicales grâce à une meilleure résilience avec une plus
haute diversité. Mais cette conclusion n'a de sens que si l'exploitation
sélective est durable \citep{Zimmerman2012}. Au contraire, une gestion
non durable des forêts tropicales entraînera des rétroactions négatives
diminuant lentement la diversité et donc la résilience des forêts,
aboutissant ultimement à la dégradation des forêts.

Forest disturbances are the third worldwide source of greenhous gas.
Tropical deforestation and degradation emit 8.26 billion of tons of
carbon dioxyde per year \citep{Pearson2017}. Deforestation has retained
much attention, but degradation from forest represents 20\% of emissions
in brazilian Amazon \citep{Asner2005}. Sustainable forest management has
been promoted as an answer to deforestation and degradation, besides
logging sustainability has been questionned \citep{Zimmerman2012}. On
the other hand, tropical forest host over half of the Earth's
biodiversity \citep{Scheffers2012}. Consequently, we decided to study
the role of biodiversity in forest ecosystem answer to disturbance,
linking diversity to ecosystem functionning \citep{Loreau2010}. We used
the hypothesis that when a disturbance event happen, due to a higher
productivity, a more diverse forest will be more resilient, based on the
positive relationship between biodiversity and productivity. We linked
that hypothesis to the complementarity and selection effects
\citep{Loreau2001}. Complementarity is the addition of ressource
partitionning and facilitation, whereas selection effect is the result
of competitive selection. We thus focused on mechanisms involved in the
relationship between biodiversity and forest ecosystem resilience with a
simulation approach to assess long term processes. We used TROLL model
\citep{Li} to simulate 60 matures forests with growing taxonomic and
functional diversities. We disturbed all forests and measured the
resilience of their ecosystem functions. Additionnally, we measured
biodiversity net effect resilience partitioned into complementarity and
selection effects. We found that diversity improved tropical forest
resilience, particularly through functional diversity and eveness.
Moreover, we showed that complementarity between species insured forest
recovery in the beginning of the succession before being replaced by
selection effect. Our results suggest the possibility for a sustainable
management of tropical forest due to an increased resilience with an
higher diversity. But this conclusion has meaning only if selective
logging meet sustainability \citep{Zimmerman2012}. On the contrary,
unsustainable tropical forest management will lead to negative feedbacks
slowly diminishing diversity and thus forest resilience, resulting
ultimately in forest degradation.

@@

\begin{itemize}
\tightlist
\item
  Bruno
\item
  Stéphane
\item
  Camilla
\item
  Fabian
\item
  Isabelle
\item
  Laurent
\item
  Jérôme
\item
  Aurélie
\item
  Éric
\item
  Maxime
\item
  Raphaël
\item
  AMAP colleagues
\item
  Myriam
\item
  BIOGECO colleagues
\item
  Pascal Pe
\item
  Paracou team
\item
  Pascal Pa
\item
  Aurélie C
\end{itemize}

@@

Tropical forests disturbances, through deforestation and degradation,
account for 8.26 billion of tons of carbon dioxyde emissions per year
\citep{Pearson2017}. It represents the third source of greenhouse gas.
Besides deforestation has been rightly the focus of worldwide attention,
degradation has been less studied and quantified through tropics.
Degradation from forest has been estimated to represents 10 time
deforestation \citep{Herold2011} and represents 20\% of greenhouse gas
emissions in the brazilian Amazon \citep{Asner2005}. Degradation has met
numerous definitions \citep{Simula2009}, but can be defined as the
result from a disturbance event that induced a modification of the
forest ecosystem reducing ecosystem services while keeping the ecosystem
as a forest, contrary to deforestation.

@@

We decided to use the forest model TROLL to study the role of
biodiversity in forest ecosystem answer to disturbance. Resilience
encompass several definitions but was summarized by \citet{Oliver2015}
as the degree of resistance or fast recovery from an ecosystem function
to environmental disturbance.

@@

In the limit of the model, we were able to show that diversity improved
tropical forest resilience. More particularly, functional diversity and
evenness are key components of diversity in forest recovery after
disturbance. Moreover, we found that complementarity between species was
insuring forest recovery in forest succession start with facilitation
before more productive species dominate the forest and insure recovery.
Our results advocates for a sustainable harvesting of tropical forests
through an increased resilience due to high diversity. But this
conclusion should met a sustainable definiton of selective logging
following \citet{Zimmerman2012}. Because if the harvesting is not
sustainable negative feedbacks will slowly diminish diversity and its
benefits for forest resilience, resulting in forest degradation.

@@

% Last pages
  %Last page
  \newpage
  \scriptsize{
  \paragraph{Résumé :}
  Les forêts tropicales font face à de nombreuses perturbations qui représentent la troisième source mondiale d’émission de gaz à effet de serre. La déforestation et la dégradation des forêts tropicales sont responsables de l’émission de 8.26 milliards de tonnes de dioxyde de carbone par an (Pearson et al. 2017). La déforestation a retenu l’attention mondiale, mais la dégradation des forêts représente 20\% des émissions de l’Amazonie brésilienne (Asner et al. 2005). La gestion durable des forêts a été proposée comme réponse à la déforestation et la dégradation, malgré la remise en question de la durabilité de l’exploitation forestière (Zimmerman \& Kormos 2012). D’autre part, les forêts tropicales abritent plus de la moitié de la biodiversité terrestre mondiale (Scheffers et al. 2012). Par conséquent, nous avons décidé d’étudier le rôle de la biodiversité dans la réponse des écosystèmes forestiers aux perturbations, en reliant diversité et fonctionnement de l’écosystème (Loreau 2010). Nous avons utilisé l’hypothèse que lors d’une perturbation, grâce à une productivité plus forte, une forêt plus diverse aura une meilleure résilience, en se basant sur la relation positive entre biodiversité et productivité. Nous avons relié cette hypothèse aux effets de complémentarité et de sélection (Loreau \& Hector 2001b). La complémentarité est la combinaison de la partition des ressources et de la facilitation, alors que l’effet de sélection est le résultat de la sélection compétitive. Nous avons ainsi centré l’étude sur les mécanismes impliqués dans la relation entre biodiversité et résilience des écosystème forestiers par une approche par simulation afin d’appréhender les processus à long terme. Nous avons utilisé le modèle TROLL (Maréchaux \& Chave) pour simuler 60 forêts matures aux diversités taxonomiques et fonctionnelles croissantes. Nous avons perturbé toutes les forêts et mesuré la résilience de leurs fonctions écosystémiques. En outre, nous avons mesuré la résilience de l’effet net de la biodiversité que l’on a décomposé en en effets de complémentarité et de sélection. Nous avons trouvé que la diversité améliore la résilience des forêts tropicales, particulièrement au travers de la diversité et l’équitabilité fonctionnelle. De plus, nous avons montré que la complémentarité entre les espèces assurait la résilience de la forêt en début de succession avant de laisser place à l’effet de sélection. Nos résultats suggèrent la possibilité d’une gestion durable des forêts tropicales grâce à une meilleure résilience avec une plus haute diversité. Mais cette conclusion n’a de sens que si l’exploitation sélective est durable (Zimmerman \& Kormos 2012). Au contraire, une gestion non durable des forêts tropicales entraînera des rétroactions négatives diminuant lentement la diversité et donc la résilience des forêts, aboutissant ultimement à la dégradation des forêts.
  \paragraph{Mots clés :} Résilience, Biodiveristé, Exploitation sélective, Fonctionnement de l'écosystème
  \paragraph{Abstract:}
  Forest disturbances are the third worldwide source of greenhous gas. Tropical deforestation and degradation emit 8.26 billion of tons of carbon dioxyde per year (Pearson et al. 2017). Deforestation has retained much attention, but degradation from forest represents 20\% of emissions in brazilian Amazon (Asner et al. 2005). Sustainable forest management has been promoted as an answer to deforestation and degradation, besides logging sustainability has been questionned (Zimmerman \& Kormos 2012). On the other hand, tropical forest host over half of the Earth’s biodiversity (Scheffers et al. 2012). Consequently, we decided to study the role of biodiversity in forest ecosystem answer to disturbance, linking diversity to ecosystem functionning (Loreau 2010). We used the hypothesis that when a disturbance event happen, due to a higher productivity, a more diverse forest will be more resilient, based on the positive relationship between biodiversity and productivity. We linked that hypothesis to the complementarity and selection effects (Loreau \& Hector 2001b). Complementarity is the addition of ressource partitionning and facilitation, whereas selection effect is the result of competitive selection. We thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience with a simulation approach to assess long term processes. We used TROLL model (Maréchaux \& Chave) to simulate 60 matures forests with growing taxonomic and functional diversities. We disturbed all forests and measured the resilience of their ecosystem functions. Additionnally, we measured biodiversity net effect resilience partitioned into complementarity and selection effects. We found that diversity improved tropical forest resilience, particularly through functional diversity and eveness. Moreover, we showed that complementarity between species insured forest recovery in the beginning of the succession before being replaced by selection effect. Our results suggest the possibility for a sustainable management of tropical forest due to an increased resilience with an higher diversity. But this conclusion has meaning only if selective logging meet sustainability (Zimmerman \& Kormos 2012). On the contrary, unsustainable tropical forest management will lead to negative feedbacks slowly diminishing diversity and thus forest resilience, resulting ultimately in forest degradation.
  \paragraph{Keywords:}Resilience, Biodiversity, Selective logging, Ecosystem functionning
  }

### 2017-07-19 latex/after_body.tex
Abstract 1

%Last page
\newpage
\paragraph{Résumé :}
Écrire le résumé ici...
\paragraph{Mots clés :} mots clés
\newline\newline
\paragraph{Abstract:}
Write abstract here
\paragraph{Keywords:} keywords

>>>>

%Last page
\newpage
\scriptsize{
\paragraph{Résumé :}
Les forêts tropicales font face à de nombreuses perturbations qui représentent la troisième source mondiale d’émission de gaz à effet de serre. La déforestation et la dégradation des forêts tropicales sont responsables de l’émission de 8.26 milliards de tonnes de dioxyde de carbone par an (Pearson et al. 2017). La déforestation a retenu l’attention mondiale, mais la dégradation des forêts représente 20\% des émissions de l’Amazonie brésilienne (Asner et al. 2005). La gestion durable des forêts a été proposée comme réponse à la déforestation et la dégradation, malgré la remise en question de la durabilité de l’exploitation forestière (Zimmerman \& Kormos 2012). D’autre part, les forêts tropicales abritent plus de la moitié de la biodiversité terrestre mondiale (Scheffers et al. 2012). Par conséquent, nous avons décidé d’étudier le rôle de la biodiversité dans la réponse des écosystèmes forestiers aux perturbations, en reliant diversité et fonctionnement de l’écosystème (Loreau 2010). Nous avons utilisé l’hypothèse que lors d’une perturbation, grâce à une productivité plus forte, une forêt plus diverse aura une meilleure résilience, en se basant sur la relation positive entre biodiversité et productivité. Nous avons relié cette hypothèse aux effets de complémentarité et de sélection (Loreau \& Hector 2001b). La complémentarité est la combinaison de la partition des ressources et de la facilitation, alors que l’effet de sélection est le résultat de la sélection compétitive. Nous avons ainsi centré l’étude sur les mécanismes impliqués dans la relation entre biodiversité et résilience des écosystème forestiers par une approche par simulation afin d’appréhender les processus à long terme. Nous avons utilisé le modèle TROLL (Maréchaux \& Chave) pour simuler 60 forêts matures aux diversités taxonomiques et fonctionnelles croissantes. Nous avons perturbé toutes les forêts et mesuré la résilience de leurs fonctions écosystémiques. En outre, nous avons mesuré la résilience de l’effet net de la biodiversité que l’on a décomposé en en effets de complémentarité et de sélection. Nous avons trouvé que la diversité améliore la résilience des forêts tropicales, particulièrement au travers de la diversité et l’équitabilité fonctionnelle. De plus, nous avons montré que la complémentarité entre les espèces assurait la résilience de la forêt en début de succession avant de laisser place à l’effet de sélection. Nos résultats suggèrent la possibilité d’une gestion durable des forêts tropicales grâce à une meilleure résilience avec une plus haute diversité. Mais cette conclusion n’a de sens que si l’exploitation sélective est durable (Zimmerman \& Kormos 2012). Au contraire, une gestion non durable des forêts tropicales entraînera des rétroactions négatives diminuant lentement la diversité et donc la résilience des forêts, aboutissant ultimement à la dégradation des forêts.
\paragraph{Mots clés :} Résilience, Biodiveristé, Exploitation sélective, Fonctionnement de l'écosystème
\paragraph{Abstract:}
Forest disturbances are the third worldwide source of greenhous gas. Tropical deforestation and degradation emit 8.26 billion of tons of carbon dioxyde per year (Pearson et al. 2017). Deforestation has retained much attention, but degradation from forest represents 20\% of emissions in brazilian Amazon (Asner et al. 2005). Sustainable forest management has been promoted as an answer to deforestation and degradation, besides logging sustainability has been questionned (Zimmerman \& Kormos 2012). On the other hand, tropical forest host over half of the Earth’s biodiversity (Scheffers et al. 2012). Consequently, we decided to study the role of biodiversity in forest ecosystem answer to disturbance, linking diversity to ecosystem functionning (Loreau 2010). We used the hypothesis that when a disturbance event happen, due to a higher productivity, a more diverse forest will be more resilient, based on the positive relationship between biodiversity and productivity. We linked that hypothesis to the complementarity and selection effects (Loreau \& Hector 2001b). Complementarity is the addition of ressource partitionning and facilitation, whereas selection effect is the result of competitive selection. We thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience with a simulation approach to assess long term processes. We used TROLL model (Maréchaux \& Chave) to simulate 60 matures forests with growing taxonomic and functional diversities. We disturbed all forests and measured the resilience of their ecosystem functions. Additionnally, we measured biodiversity net effect resilience partitioned into complementarity and selection effects. We found that diversity improved tropical forest resilience, particularly through functional diversity and eveness. Moreover, we showed that complementarity between species insured forest recovery in the beginning of the succession before being replaced by selection effect. Our results suggest the possibility for a sustainable management of tropical forest due to an increased resilience with an higher diversity. But this conclusion has meaning only if selective logging meet sustainability (Zimmerman \& Kormos 2012). On the contrary, unsustainable tropical forest management will lead to negative feedbacks slowly diminishing diversity and thus forest resilience, resulting ultimately in forest degradation.
\paragraph{Keywords:}Resilience, Biodiversity, Selective logging, Ecosystem functionning
}
